The lattice of acyclic pipe dreams

### 2022-08-25 acyclicFacets/draft/latticePipeDreams.tex
Fixing my email and web address (just making simple commit to be registered)
Added Section 6 about subword complexes

\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, a4wide, pdflscape, multido, xargs, colortbl, multicol, multirow, calc, shuffle}
\hypersetup{colorlinks=true, citecolor=PineGreen, linkcolor=PineGreen}
%\usepackage[all]{xy}
\usepackage{tikz}\usetikzlibrary{trees,snakes,shapes,arrows,matrix,calc}
\usepackage{comment}
\usepackage{etex}
%\reserveinserts{50}
\graphicspath{{figures/}}
\makeatletter
\def\input@path{{figures/}}
\makeatother

@@

\author[N.~Bergeron]{Nantel Bergeron} 
\address[N.~Bergeron]{Department of Mathematics and Statistics, York University, Toronto}
\email{bergeron@mathstat.yorku.ca}
\urladdr{http://www.math.yorku.ca/bergeron/}

@@

% theorems
\newtheorem{theorem}{Theorem}%[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{question}[theorem]{Question}

@@

% operators
\DeclareMathOperator{\conv}{conv} % convex hull
\DeclareMathOperator{\cone}{cone} % cone hull
\DeclareMathOperator{\arr}{Arr} % arrangements
\DeclareMathOperator{\inv}{inv} % inversion set

% others
\newcommand{\fix}[1]{{\bf FIXME: }#1} % emphasis of a problem to FIX
\newcommand{\fref}[1]{Figure~\ref{#1}} % reference figures
\newcommand{\ie}{\textit{i.e.}~} % id est
\newcommand{\eg}{\textit{e.g.}~} % exempli gratia
\newcommand{\Eg}{\textit{E.g.}~} % exempli gratia
\newcommand{\aka}{\textit{aka.}~} % also known as
\newcommand{\viceversa}{\textit{vice versa}} % vice versa
\newcommand{\ordinal}{\textsuperscript{th}} % th for ordinals
\newcommand{\ex}[1]{^{\textrm{ex#1}}} % example
\newcommand{\para}[1]{\medskip\noindent\textbf{#1~---}} % paragraph
\newcommand{\subpara}[1]{\smallskip\noindent\textit{#1.}} % paragraph
\definecolor{PineGreen}{RGB}{2,120,120} % pinegreen color
\definecolor{darkgreen}{RGB}{57,181,74} % darkgreen color
\renewcommand{\b}[1]{{\color{blue} #1}} % blue
\renewcommand{\r}[1]{{\color{red} #1}} % red
\newcommand{\g}[1]{{\color{darkgreen} #1}} % green
\newcommand{\defn}[1]{\textbf{\textsf{\color{PineGreen} #1}}} % emphasis of a definition
\usepackage{todonotes}
\newcommand{\nantel}[1]{\todo[color=red!30]{#1 \\ \hfill --- N.}}
\newcommand{\cesar}[1]{\todo[color=orange!30]{#1 \\ \hfill --- C.}}
\newcommand{\noemie}[1]{\todo[color=green!30]{#1 \\ \hfill --- N.}}
\newcommand{\vincent}[1]{\todo[color=blue!30]{#1 \\ \hfill --- V.}}

@@

% pipe dreams
\input{pipeDreamsTikz}
% cross
%\newcommand{\cross}[1][black]{\raisebox{.1cm}{\pipeDreamMonoColor{#1}{black}{c}}}
\newcommand{\cross}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{cross}}}
\newcommand{\crossBiColor}[2]{\raisebox{.1cm}{\pipeDreamBiColor{#1}{#2}{black}{c/l/r}}}
\newcommand{\NScross}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{NScross}}}
\newcommand{\WEcross}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{WEcross}}}
% elbow
%\newcommand{\elbow}[1][black]{\raisebox{.1cm}{\pipeDreamMonoColor{#1}{black}{e}}}
\newcommand{\elbow}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{elbow}}}
\newcommand{\elbowBiColor}[2]{\raisebox{.1cm}{\pipeDreamBiColor{#1}{#2}{black}{e/l/r}}}
%\newcommand{\SEelbow}[1][black]{\raisebox{.1cm}{\pipeDreamBiColor{white}{#1}{black}{e/l/r}}}
\newcommand{\SEelbow}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{SEelbow}}}
%\newcommand{\WNelbow}[1][black]{\raisebox{.1cm}{\pipeDreamBiColor{#1}{white}{black}{e/l/r}}}
\newcommand{\WNelbow}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{WNelbow}}}
\newcommand{\waving}{\mathsf{wav}}%{\raisebox{-.2cm}{\begin{tikzpicture}[scale=1]\pipeDreamBiColor{.2}{(0,0)}{black}{white}{{t/r/l},{t/l/l},{t/l/r}}\end{tikzpicture}}}
\newcommand{\transpose}{\mathsf{transp}}
\newcommand{\pipeDreams}{\Pi} % set of pipe dreams
\newcommand{\HP}{{\bk\pipeDreams}} % Hopf Algebra of pipe dreams
\newcommand{\reversingPipeDreams}{\Omega} % reversing pipe dreams
\newcommand{\contact}{^\#} % contact graph
\newcommand{\se}{\textsc{se}} % south-east
\newcommand{\nw}{\textsc{nw}} % north-west
\newcommand{\insertion}[2]{\mathsf{ins}(#1,#2)} % insertion of permutation #1 in pipe dreams with permutation #2
\newcommand{\duality}{^\star} % duality pipe dreams - triangulations
\newcommand{\acyclicPipeDreams}{\Sigma} % acyclic pipe dreams
\newcommand{\wo}{w_\circ} % longest element
\newcommand{\subwordComplex}{\mathcal{SC}} % subword complex
\newcommand{\Roots}{\mathrm{R}} % root configuration
\newcommand{\linearExtensions}{\mathcal{L}} % linear extensions
\newcommand{\noninversions}[2]{\mathsf{ninv}(#1,#2)} % non-inversions
\newcommand{\acyclicOrientations}{\Omega} % set of pipe dreams
\newcommand{\recoils}[2]{\mathsf{rec}(#1,#2)} % recoil scheme of a pipe dream
\newcommand{\canopy}[1]{\mathsf{can}(#1)} % canopy of a pipe dream

@@

\begin{theorem}
For any permutation~$\omega \in \fS_n$, the pipe dream congruence~$\equiv_\omega$ is a lattice congruence of the interval~$[e,\omega]$ of the weak order.
\end{theorem}

@@

\begin{corollary}
For any permutation~$\omega \in \fS_n$, the increasing flip poset on~$\acyclicPipeDreams(\omega)$ is a lattice.
\end{corollary}

@@

>>>>

\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, a4wide, pdflscape, multido, xargs, colortbl, multicol, multirow, calc, shuffle}
\hypersetup{colorlinks=true, citecolor=PineGreen, linkcolor=PineGreen}
%\usepackage[all]{xy}
\usepackage{tikz}\usetikzlibrary{trees,snakes,shapes,arrows,matrix,calc}
\usepackage{comment}
\usepackage{etex}
\usepackage{ulem} % to strike through a word

%\reserveinserts{50}
\graphicspath{{figures/}}
\makeatletter
\def\input@path{{figures/}}
\makeatother

@@

\author[N.~Bergeron]{Nantel Bergeron} 
\address[N.~Bergeron]{Department of Mathematics and Statistics, York University, Toronto}
\email{bergeron@yorku.ca}
\urladdr{http://bergeron.mathstats.yorku.ca}

@@

% theorems
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{question}[theorem]{Question}

@@

% operators
\DeclareMathOperator{\conv}{conv} % convex hull
\DeclareMathOperator{\cone}{cone} % cone hull
\DeclareMathOperator{\arr}{Arr} % arrangements
\DeclareMathOperator{\Inv}{Inv} % inversion set
\DeclareMathOperator{\Ninv}{Ninv} % non-inversion set
\DeclareMathOperator{\DemazureProduct}{Dem} % non-inversion set

% others
\newcommand{\fix}[1]{{\bf FIXME: }#1} % emphasis of a problem to FIX
\newcommand{\fref}[1]{Figure~\ref{#1}} % reference figures
\newcommand{\ie}{\textit{i.e.}~} % id est
\newcommand{\eg}{\textit{e.g.}~} % exempli gratia
\newcommand{\Eg}{\textit{E.g.}~} % exempli gratia
\newcommand{\aka}{\textit{aka.}~} % also known as
\newcommand{\viceversa}{\textit{vice versa}} % vice versa
\newcommand{\ordinal}{\textsuperscript{th}} % th for ordinals
\newcommand{\ex}[1]{^{\textrm{ex#1}}} % example
\newcommand{\para}[1]{\medskip\noindent\textbf{#1~---}} % paragraph
\newcommand{\subpara}[1]{\smallskip\noindent\textit{#1.}} % paragraph
\definecolor{PineGreen}{RGB}{2,120,120} % pinegreen color
\definecolor{darkgreen}{RGB}{57,181,74} % darkgreen color
\renewcommand{\b}[1]{{\color{blue} #1}} % blue
\renewcommand{\r}[1]{{\color{red} #1}} % red
\newcommand{\g}[1]{{\color{darkgreen} #1}} % green
\newcommand{\defn}[1]{\textbf{\textsf{\color{PineGreen} #1}}} % emphasis of a definition
\usepackage{todonotes}
\newcommand{\nantel}[1]{\todo[color=red!30]{#1 \\ \hfill --- N.}}
\newcommand{\cesar}[1]{\todo[color=orange!30,inline]{#1 \\ \hfill --- C.}}
\newcommand{\cesarm}[1]{\todo[color=orange!30]{#1 \\ \hfill --- C.}}
\newcommand{\noemie}[1]{\todo[color=green!30]{#1 \\ \hfill --- N.}}
\newcommand{\vincent}[1]{\todo[color=blue!30]{#1 \\ \hfill --- V.}}

@@

% pipe dreams
\input{pipeDreamsTikz}
% cross
%\newcommand{\cross}[1][black]{\raisebox{.1cm}{\pipeDreamMonoColor{#1}{black}{c}}}
\newcommand{\cross}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{cross}}}
\newcommand{\crossBiColor}[2]{\raisebox{.1cm}{\pipeDreamBiColor{#1}{#2}{black}{c/l/r}}}
\newcommand{\NScross}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{NScross}}}
\newcommand{\WEcross}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{WEcross}}}
% elbow
%\newcommand{\elbow}[1][black]{\raisebox{.1cm}{\pipeDreamMonoColor{#1}{black}{e}}}
\newcommand{\elbow}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{elbow}}}
\newcommand{\elbowBiColor}[2]{\raisebox{.1cm}{\pipeDreamBiColor{#1}{#2}{black}{e/l/r}}}
%\newcommand{\SEelbow}[1][black]{\raisebox{.1cm}{\pipeDreamBiColor{white}{#1}{black}{e/l/r}}}
\newcommand{\SEelbow}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{SEelbow}}}
%\newcommand{\WNelbow}[1][black]{\raisebox{.1cm}{\pipeDreamBiColor{#1}{white}{black}{e/l/r}}}
\newcommand{\WNelbow}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{WNelbow}}}
\newcommand{\waving}{\mathsf{wav}}%{\raisebox{-.2cm}{\begin{tikzpicture}[scale=1]\pipeDreamBiColor{.2}{(0,0)}{black}{white}{{t/r/l},{t/l/l},{t/l/r}}\end{tikzpicture}}}
\newcommand{\transpose}{\mathsf{transp}}
\newcommand{\pipeDreams}{\Pi} % set of pipe dreams
\newcommand{\HP}{{\bk\pipeDreams}} % Hopf Algebra of pipe dreams
\newcommand{\reversingPipeDreams}{\Omega} % reversing pipe dreams
\newcommand{\contact}{^\#} % contact graph
\newcommand{\se}{\textsc{se}} % south-east
\newcommand{\nw}{\textsc{nw}} % north-west
\newcommand{\insertion}[2]{\mathsf{ins}(#1,#2)} % insertion of permutation #1 in pipe dreams with permutation #2
\newcommand{\duality}{^\star} % duality pipe dreams - triangulations
\newcommand{\acyclicPipeDreams}{\Sigma} % acyclic pipe dreams
\newcommand{\linearExtensions}{\mathcal{L}} % linear extensions
\newcommand{\noninversions}[2]{\mathsf{ninv}(#1,#2)} % non-inversions
\newcommand{\acyclicOrientations}{\Omega} % set of pipe dreams
\newcommand{\recoils}[2]{\mathsf{rec}(#1,#2)} % recoil scheme of a pipe dream
\newcommand{\canopy}[1]{\mathsf{can}(#1)} % canopy of a pipe dream

% Subword complexes

\newcommand{\wo}{w_\circ} % longest element
\newcommand{\subwordComplex}{\mathcal{SC}} % subword complex
\newcommand{\Roots}{\mathrm{R}} % root configuration
\newcommand{\rootFunction}[2]{\mathrm{r}_{#1}(#2)} % root function
\newcommand{\subwordFacets}{\mathcal{F}} % subword complex
\newcommand{\subwordAcyclicFacets}{\mathcal{F}^{ac}} % subword complex
\newcommand{\greedyFacet}{I^{\overleftarrow{gr}}} % greedy facet
\newcommand{\antiGreedyFacet}{I^{\overrightarrow{gr}}} % antigreedy facet
\newcommand{\sweepingAlgorithm}{\mathrm{sweep}} % sweeping algorithm

@@

\begin{theorem}\label{thm_pipe_latticecongruence}
For any permutation~$\omega \in \fS_n$, the pipe dream congruence~$\equiv_\omega$ is a lattice congruence of the interval~$[e,\omega]$ of the weak order.
\end{theorem}

@@

\begin{corollary}\label{cor_acyclicpipedreams_lattice}
For any permutation~$\omega \in \fS_n$:
\begin{enumerate}
\item The increasing flip poset on~$\acyclicPipeDreams(\omega)$ is isomorphic to the lattice quotient~$[e,\omega]/\equiv_\omega$.
\item The increasing flip poset on~$\acyclicPipeDreams(\omega)$ is a lattice.
\end{enumerate}
\end{corollary}

@@

\newpage
\section{Subword complexes}
The objective of this section is to partially extend the results about the lattice structure of acyclic pipe dreams, presented in Section~\ref{sec:latticeAcyclicPipeDreams}, to the wider context of subword complexes of finite type. Before doing so, we recall some preliminaries about finite Coxeter groups and subword complexes. We refer to~\cite{bjorner_coxeterGroups_2005,humphreys_reflectionGroups_1990} for more details on Coxeter groups.

\subsection{Finite coxeter groups} 
%a~\cesar{To do:\\
%Coxeter groups of finite type, root systems, weak order, Bruhat order, inversions, non-inversions. 
%}
We consider a \defn{finite root system} $\Phi=\Phi^+\cup \Phi^-$ in an $n$-dimensional euclidian space $V$, with \defn{simple roots} $\Delta=\{\alpha_1,\dots,\alpha_n\} \subseteq \Phi^+$. 
The reflections along the hyperplanes orthogonal to the roots in $\Phi$ generate a finite Coxeter group $W$.  
This group is generated by the simple transpositions $S=\{s_1,\dots ,s_n\}$, where $s_i$ is the reflection along the hyperplane orthogonal to the simple root $\alpha_i$. We also denote by $\alpha_s$ the simple root associated to $s\in S$. Throughout this paper, we only consider Coxeter groups which are finite. They are classified in terms of Dynkin diagrams, and arise from finite root systems, see~\cite{humphreys_reflectionGroups_1990}.

Abstractly, the \defn{Coxeter group} $W$ is the group generated by a collection $S=\{s_1,\dots, s_n\}$ which satisfies the relations $(s_is_j)^{m_{ij}}=1$, where $m_{ii}=1$ and $m_{ij}\geq 2$ for~$i\neq j$.
The pair $(W,S)$ is called a \defn{Coxeter system}.
The \defn{length} $\ell(w)$ of an element $w\in W$ is the smallest~$\ell$ such that it can be written of the form $w=w_1w_2\dots w_\ell$ where $w_i\in S$. A \defn{reduced expression} of~$w$ is an expression of minimal length. 
We denote by $\wo$ the \defn{unique longest} element in $W$.

For a finite Coxeter $W$ associated to a root system $\Phi$, the length $\ell(w)$ can be geometrically interpreted as the cardinality of the \defn{inversion set} $\Inv(w)$ of $w$, which is defined as
\begin{align}\label{def_inversionset}
\Inv(w) := \Phi^+ \cap w(\Phi^-). 
\end{align}
Indeed, this set can be written as 
\begin{align}\label{def_inversionset2}
\Inv(w) := \{
\alpha_{w_1}, w_1(\alpha_{w_2}),\dots , w_1w_2\dots w_{\ell-1}(\alpha_{w_\ell})
\}
\end{align}
for any reduced expression $w=w_1w_2\dots w_\ell$ of $w$.

The \defn{non-inversion set} $\Ninv (w)$ of $w$ is defined as
\begin{align}
\Ninv(w) &:= \Phi^+ \cap w(\Phi^+) \label{def_noninversionset} \\
& =\Inv(w\wo)
\end{align}

In particular, 
\begin{align}
\Phi^+=\Inv(w) \sqcup \Ninv (w).
\end{align}

The \defn{weak order} on $W$ is the partial order $\leq$ defined by $u \leq w$ if there exists $v \in W$ such that $uv = w$ and $\ell(u) + \ell(v) = \ell(w)$. In other words, the element $w$ has a reduced expression with a prefix which is a reduced expression of $u$. Equivalently, the weak order corresponds to the inclusion order on inversion sets, that is $u \leq w$ if and only if $\Inv(u) \subseteq \Inv(w)$. This order defines a lattice structure on the elements of W. The minimal element is the identity $e \in W$ and the maximal element is $\wo$.

\subsection{Subword complexes}
%~\cesar{To do:\\
%subword complexes, flips, greedy facets, \sout{root function, root configuration, acyclic facets}. 
%}

Motivated by their study of Gröbner geometry of Schubert varieties~\cite{KnutsonMiller-GroebnerGeometry}, Knutson and Miller~\cite{knutson_subword_2004} introduced a remarkable family of simplicial complexes in the context of Coxeter groups.

Let $(W,S)$ be a Coxeter system of finite type, $Q=(q_1,\dots,q_m)$ be a word in the generators $S$, and $w\in W$ be an element of the group. 
For $J\subseteq [r]$, we denote by $Q_J$ the subword of~$Q$ consisting of the letters with positions in $J$. 
The \defn{subword complex} $\subwordComplex(Q,w)$ is the simplicial complex whose facets are subsets $I\subseteq [m]$ such that $Q_{[m]\setminus I}$ is a reduced expression for $w$. 
We denote the \defn{set of facets} of $\subwordComplex(Q,w)$ by $\subwordFacets(Q,w)$.

Two facets $I,J\in\subwordComplex(Q,w)$ are said to be connected by a \defn{flip} if they differ one element:
\[J=I\smallsetminus \{i\} \cup\{j\}\]
for some $i\in I$ and $j\in J$ with $i\neq j$. If $i<j$, the flip from $I$ to $J$ is called an \defn{increasing flip}.

The increasing flip graph on facets of the subword complex is an acyclic graph which has a unique source and a unique sink~\cite{pilaud_greedy_2012,pilaud_el_labelling_2013}.
These two special facets are called the \defn{greedy facet} $\greedyFacet$ and the \defn{antigreedy facet} $\antiGreedyFacet$. They are the lexicographically smallest and largest facets of $\subwordComplex(Q,\pi)$, respectively.

The poset on the set of facets induced by increasing flips is called the \defn{increasing flip poset}.

\begin{definition}[{\cite{CeballosLabbeStump}}]
    The \defn{root function} $\rootFunction{I}{\cdot}:[m]\rightarrow \Phi$ associated to a facet $I$ is the map from positions in the word $Q$ to roots defined by 
    \[
    \rootFunction{I}{k} := \prod Q_{[k-1]\smallsetminus I}(\alpha_{q_k}).
    \]
    The \defn{root configuration} is the set $\Roots(I)=\set{\rootFunction{I}{i}}{i\in I}$. 
    The facet $I$ is called \defn{acyclic} if $\cone \Roots(I)$ is a pointed cone. 
    We denote the \defn{set of acyclic facets} by $\subwordAcyclicFacets(Q,w)$. 
\end{definition}
The root function is a useful and well established tool in the study of subword complexes. For instance, they can be used to fully understand flips. 
\begin{lemma}[{\cite{CeballosLabbeStump,knutson_subword_2004}}]
\label{lem_rootfunction_flips}
Let $I$ be a facet of the subword complex~$\subwordComplex(Q,w)$. The following hold:
\begin{enumerate}
    \item $i \in I$ is flippable if and only if $|\rootFunction{I}{i}|\in \Inv(w)$.
    \item If $i \in I$ is flippable, it can be flipped to the unique $j\in [m]\smallsetminus I$ such that $\rootFunction{I}{j}=|\rootFunction{I}{i}|$. 
    The flip is increasing ($i<j$) when $\rootFunction{I}{i}\in \Phi^+$ and decreasing ($i>j$) when $\rootFunction{I}{i}\in \Phi^-$. \label{lem_rootfunction_flips2}
    \item If $I'=I\smallsetminus \{i\}\cap \{i'\}$ is a facet obtained from $I$ by flipping $i$ to $i'$ with $i<i'$, then
    \begin{equation}\label{eq_rootfunction_flip}
         \rootFunction{I'}{k} = 
  \begin{cases}
    s_\beta(\rootFunction{I}{k}), & \text{for } i < k \leq i' \\
    \rootFunction{I}{k}, & \text{otherwise}
  \end{cases}
    \end{equation}
    where $\beta=\rootFunction{I}{i}$ and $s_\beta\in W$ is the reflection orthogonal to the root $\beta$.
    \item $i \in I$ is not flippable if and only if $\rootFunction{I}{i}\in \Ninv(w)$.
\end{enumerate}
\end{lemma}

\subsection{Linear extensions of facets}

\begin{definition}
Let $\subwordComplex(Q,w)$ be a non-empty subword complex and $I\in \subwordComplex(Q,w)$ be a facet. 
An element $\pi\in W$ is called a \defn{linear extension} of $I$ if the root configuration $\Roots(I)\subseteq \pi(\Phi^+)$.
The \defn{set of linear extensions} of $I$ is denoted by
\begin{align}
\linearExtensions(I)=
\set{\pi\in W}{\Roots(I)\subseteq \pi(\Phi^+)}.    
\end{align}
\end{definition}

\begin{lemma}
\label{lem_acyclic_nonempty}
$\linearExtensions(I)\neq \emptyset$ if and only if $I$ is acyclic.
\end{lemma}
\begin{proof}
Let $\Phi=\Phi^+\sqcup\Phi^-$ be the root system associated to $W$. 
A known result for root systems of finite type states that any other choice of a positive root system $\Phi^+$ for $\Phi$ is of the form $\widetilde \Phi^+=\pi(\Phi^+)$ for some $w\in W$. 
Since any pointed cone is contained in some $\widetilde \Phi^+$, then 
\[
\begin{array}{rcl}
I \text{ is acyclic}  & \longleftrightarrow  &  \cone \Roots(I) \text{ is pointed} \\
  & \longleftrightarrow  & \cone \Roots(I)\subseteq \pi(\Phi^+) \text{ for some } \pi\in W \\
  & \longleftrightarrow  &  \Roots(I)\subseteq \pi(\Phi^+) \text{ for some } \pi\in W \\
  & \longleftrightarrow  &  \linearExtensions(I) \neq \emptyset 
\end{array}
\]
\end{proof}

\begin{lemma}
\label{lem_pi_phiplus_inv_ninv}
The following holds,
\[
\pi(\Phi^+)=-\Inv(\pi)\sqcup \Ninv(\pi).
\]
\end{lemma}
\begin{proof}
This is straight forward from the definition of inversions and non-inversion of $\pi$. In fact, multiplying Equation~\eqref{def_inversionset} by $-1$ we deduce $\pi(\Phi^+)\cap \Phi^-=-\Inv(\pi)$. 
Furthermore,  
$\pi(\Phi^+)\cap \Phi^+=\Ninv(\pi)$ by Equation~\eqref{def_noninversionset}.
\end{proof}

The following lemma regards linear extensions of the greedy and antigreedy facets.

\begin{lemma}\label{lem_linearext_greedyfacets}
    Let $\greedyFacet$ and $\antiGreedyFacet$ be the greedy and antigreedy facets of $\subwordComplex(Q,w)$, respectively. Then 
    \begin{enumerate}
        \item $e\in \linearExtensions(\greedyFacet)$. 
        \item $w \in \linearExtensions(\antiGreedyFacet)$.
        \item If $e\in \linearExtensions(I)$ then $I=\greedyFacet$.
    \end{enumerate}
\end{lemma}
\begin{proof}
    Let $e\in W$ be the minimal element of the week order. 
    The greedy facet $\greedyFacet$ is the unique facet for which every flip (if any) is increasing. By Lemma~\ref{lem_rootfunction_flips}~\eqref{lem_rootfunction_flips2}, this implies that $\greedyFacet$ is the unique facet $I$ such that $\Roots(I)\subseteq \Phi^+ = e(\Phi^+)$. This proves parts (1) and (3) of the Lemma.

    For part (2) we need to analyze the possibilities for the set $\Roots(\antiGreedyFacet)$.  By Lemma~\ref{lem_rootfunction_flips}~\eqref{lem_rootfunction_flips2}, if $i\in \antiGreedyFacet$ is flippable then $\rootFunction{\antiGreedyFacet}{i}\in -\Inv(w)$. Furthermore, if $i\in \antiGreedyFacet$ is not flippable then $\rootFunction{\antiGreedyFacet}{i}\in \Ninv(w)$. Therefore, $\Roots(\antiGreedyFacet)\subseteq w(\Phi^+)=-\Inv(w)\sqcup \Ninv(w)$.
\end{proof}

\begin{remark}
The concept of linear extensions of acyclic facets of subword complexes are a natural generalization of linear extensions of contact graphs of acyclic pipe dreams that we introduced in Section~\ref{subsec:linearExtensions}. Indeed ... 
\cesar{make this connection crystal clear...}
\end{remark}

\subsection{Two fundamental theorems about linear extensions}
In this section, we present our two main results about linear extensions for subword complexes, Theorem~\ref{thm_subword_linearextensions_A} and Theorem~\ref{thm_subword_linearextensions_B}.

We say that a subset $A$ of the weak order is a \defn{lower set} if $a<b$ and $b\in A$ implies $a\in A$. 
We say that $A$ is \defn{closed by intervals} if $a,b\in A$ implies that $[a,b]\in A$, where $[a,b]=\{c: a \leq c\leq b\}$.

\begin{theorem}\label{thm_subword_linearextensions_A}
Let~$\subwordComplex(Q,w)$ be a non-empty subword complex. The following hold:
\begin{enumerate}
\item  $\bigcup_{I\in\subwordFacets(Q,w)} \linearExtensions(I)$ is a lower set of the weak order. 
\label{thmA_item1}
\hfill (lower set)
\item $[e,w] \subseteq \bigcup_{I\in\subwordFacets(Q,w)} \linearExtensions(I)$ 
\label{thmA_item2}
\hfill (containment)
\item If $I_1\neq I_2$ then $\linearExtensions(I_1)\cap \linearExtensions(I_2)=\emptyset$. 
\label{thmA_item3}
\hfill (partition)
\item The set $\linearExtensions(I)$ is closed by intervals. 
\label{thmA_item4}
\hfill (interval closed) 
\end{enumerate}
\end{theorem}

\begin{remark}\label{rem_thmA}
Before proving this theorem, we would like to make the following remarks about items~\eqref{thmA_item2} and~\eqref{thmA_item4}.
\begin{enumerate}
  \setcounter{enumi}{1}
\item There are cases where $[e,w] \neq \bigcup_{I\in\subwordFacets(Q,w)}$. \cesarm{add discussion}
\begin{figure}[h]
\includegraphics[width=0.52\textwidth]{interval_not_contained1} \quad
\includegraphics[width=0.43\textwidth]{interval_not_contained2}
\end{figure}
  \setcounter{enumi}{3}
\item $\linearExtensions(I)$ is not an interval in general. See for instance the example~\cite[Figure~9]{PilaudStump-brickPolytope}. 
In that figure, the elements of the week order $\mathfrak{S}_4$ that are linear extensions of the same acyclic facet of a given subword complex are grouped together. Some of the groups in the figure are not intervals.  \\
\end{enumerate}
\end{remark}

\begin{proof}[Proof of Theorem~\ref{thm_subword_linearextensions_A}]
We will prove the items of the theorem one by one.

Proof of \eqref{thmA_item1}:
Let $\pi\in \linearExtensions(I)$ for some facet $I$.
We need to show that if $\pi'<\pi$ then there exist another facet $I'$ such that $\pi'\in \linearExtensions(I')$. 
It is enough to show this when $\pi=\pi' s$ for some $s\in S$ such that $\ell(\pi'
)<\ell(\pi)$.

By definition, $\pi\in \linearExtensions(I)$ if and only if $\Roots(I)\subseteq \pi(\Phi^+)$. Now, 
\[
\pi(\Phi^+)= -\Inv(\pi)\sqcup \Ninv(\pi)
\]  
and 
\begin{align*}
    \Inv(\pi') &=\Inv(\pi)\smallsetminus \{\beta\} \\
    \Ninv(\pi') &= \Ninv(\pi) \cup \{\beta\}
\end{align*}
for $\beta=\pi'(\alpha_s)$. 
Therefore,
\begin{align*}
    \pi'(\Phi^+) = \pi(\Phi^+) \smallsetminus \{-\beta\} \cup \{\beta\}.
\end{align*}

\begin{center}
\begin{tikzpicture}
\coordinate (a) at (0,0);
\coordinate (b) at (0:2);
\coordinate (c) at (60:2);
\coordinate (d) at (120:2);
\coordinate (e) at (180:2);
\draw (b) node[right] {$\beta$};
\draw (e) node[left] {$-\beta$};
%
\path[fill=green!20] (0,0) --  (60:2) arc(60:180:2) -- cycle;
%
\draw[-stealth,dashed] (a) -- (b);
\draw[-stealth] (a) -- (c);
% \draw[-stealth] (a) -- (d);
\draw[-stealth] (a) -- (e);
\node at (120:1) {$\pi(\Phi^+)$};
\end{tikzpicture} 
%
\qquad
%
\begin{tikzpicture}
\coordinate (a) at (0,0);
\coordinate (b) at (0:2);
\coordinate (c) at (60:2);
\coordinate (d) at (120:2);
\coordinate (e) at (180:2);
\draw (b) node[right] {$\beta$};
\draw (e) node[left] {$-\beta$};
%
\path[fill=green!20] (0,0) -- (0:2) arc(0:120:2)  -- cycle;
%
\draw[-stealth] (a) -- (b);
% \draw[-stealth] (a) -- (c);
\draw[-stealth] (a) -- (d);
\draw[-stealth,dashed] (a) -- (e);
\node at (60:1) {$\pi'(\Phi^+)$};
\end{tikzpicture} 
\end{center}
%
Note that the reflection $s_\beta=\pi's\pi'^{-1}$ orthogonal to the root $\beta$ transforms $\pi(\Phi^+)$ and $\pi'(\Phi^+)$ into each other:
\[
s_\beta(\pi(\Phi^+))=\pi'(\Phi^+).
\]
Since $s_\beta(\beta)= -\beta$, then $s_\beta$ preserves the set 
\[
\pi(\Phi^+)\smallsetminus \{-\beta\}=
\pi'(\Phi^+)\smallsetminus \{\beta\}=
\pi(\Phi^+)\cap \pi'(\Phi^+). 
\]
This observation will be useful to prove the statement in the second of the following two cases.

\medskip
{\bf Case 1}: $-\beta \notin \Roots(I)$.

In this case, 
$\Roots(I)\subseteq \pi(\Phi^+) \smallsetminus \{-\beta\} \subseteq \pi'(\Phi^+)$.
Taking $I'=I$, we have 
$\pi'\in \linearExtensions(I')$ 
as wanted.

\medskip
{\bf Case 2}: $-\beta \in \Roots(I)$.

In this case, we need to remove $-\beta$ from the root configuration. We will achieve this by flipping the position of the last $-\beta$ in $I$ to create a new facet $I'$. This position is indeed flippable as we will argue now.

Given a facet $I$ of a subword complex~$\subwordComplex(Q,w)$ and a positive root $\beta\in \Phi^+$, the restriction of the list of roots 
\[
\rootFunction{I}{1}, \rootFunction{I}{2}, \dots , \rootFunction{I}{m}.
\]
to the set $\{\beta,-\beta\}$ is of the form
\[
\begin{array}{cccc}
  \beta, \dots , & \beta &, -\beta, \dots, &-\beta.\\
     & i && j
\end{array}
\]
The sequence of $-\beta$'s could in principle be empty and so does the sequence of $\beta$'s. But if there is a $-\beta$ in this list, then there should be at least one $\beta$ preceding it. The position $i$ of the last $\beta$ is used in the reduced expression of $w$ in the complement of $I$, that is $i\notin I$. The positions of the other $\beta$'s and $-\beta$'s all belong to~$I$, and can all be flipped to $i$ (see Lemma~\ref{lem_rootfunction_flips}~\eqref{lem_rootfunction_flips2}).    
In particular, the position $j$ of the last $-\beta$ belongs to $I$, and it can be flipped to $i$ creating a new facet $I'=I\smallsetminus \{j\} \cap \{i\}$.
